---
title: "607 Assignment 8 Dylan Gold"
format: pdf
editor: visual
---

## Approach

In this assignment we have to make use of the NYtimes api to answer a question through queries made through the api. The hardest part of this initially was coming up with a research question. I started off by testing out the api after creating an account. I tried using the article search api and found that using the filter did not really help that much for some of my initial examples. I tried restaurants and many of the articles seeming had nothing to do with restaurants.

While doing some inital setting up of the api key in R I decide to use the usethis library to adjust the environment.\
In the console I type the following to save a api key to my environment under "NYT_API_KEY". R may need to be restarted after.\

```         

library(usethis)
edit_r_environ()
```

```{r , message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
```

Now we have our key

```{r}
api_key <- Sys.getenv("NYT_API_KEY")
```

After some testing I found that the nytimes api works with this query list. Other ways of adding it as a header did not seem to work for me but here we get an example of the current hardcover-fiction books

```{r}
url <- "https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json"
response <- GET(
  url,
  query = list('api-key' = api_key) #here we list query parameters,
)
status_code(response)
```

Looking at the data in our environment from this we can see there is a lot of extra information. I will figure out how to extract the data I need at a later point because I am still not certain of my question to ask.\

```{r}
books <- content(response)
head(books$results)
```

I will probably look at the books to answer a question about comparing different types of books or I will use the article search api to look at how many articles are made each year with a certain filter and see how things have increased. For example given some of the recent news about insider trading maybe I will use insider trading as a filter for each year within the past 15 years and see how the occurrence of these articles have either decreased or increased over time. I may pick something else as the search filter.\
At this point I at least have the api set up minimally and I will work on parsing it and deciding on my specific question later.\

## Codebase

First I will decide the question I want to answer. Given I have looked at several data sets relating to energy consumption, carbon emissions and other global warming adjacent topics this semester, I will try to answer: How has coverage of climate change changed over the past 10 years in the New York Times. I spent a while trying to answer this question but realized that getting all the articles that with the keyword subject of "Global Warming" would cause me to be rate limited. I probably could have looked one day at a time at articles slowly but I did not want to continue trying to fix a solution that may have not been viable due to a rate limit.\

So the question I am interested in answering is: What publishers and authors of books are most often best sellers within the last year?
There are many different genres of best sellers. I am hopefully expecting to see that certain authors or publishers have a higher rate of best sellers, or learn that no specific authors or publishers have several best sellers in recent years.
We can start off with an API call in the books api of the NYT times.\
I already have my account and key set up from the approach.\

```{r , message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
library(gt)
api_key <- Sys.getenv("NYT_API_KEY")

```

There are 2 calls we can make in this API a specific date and the specific best seller genre, or an overview which just takes a date and shows all the best seller genres.
I will be using the overview URL, which just takes a year and month to give all the best sellers (every genre) for I will start off with trying to get a specific year and month into a data frame format that is suitable.

```{r}
url <- "https://api.nytimes.com/svc/books/v3/lists/overview.json"
response <- GET(
  url,
  query = list( # We can put parameters here, api-key is needed and we can specify a date.
    'published_date' = '2025-01-01',
    'api-key' = api_key
    ),
)
status_code(response)
```
We have a good response code, we can try to extract the data. I can put the results into a data frame and start to inspect how to get the data we need.
```{r}
data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
json_results <- fromJSON(data, flatten = TRUE) # 
```

We have the json data of the response, we can extract the results. The results have many values describng the dates and a list containing information about each of the best seller collections. I will extract this data. Note that books is a lists type
```{r}
books_df <- json_results$results$lists
books_df
names(book_df)
```
We can see there there are 18 rows containing each of the different best seller categories. Each of the rows have a books column containing the books. I will take a closer look at this. I will look at the first row's book column.
```{r}
book1 <- book_df$books[1]
head(book1)
```
We have a data frame that contains information about each book in this list. I will select the rows we want just to visuallize it better. I also notice that the price is defaulting to 0. If the price had shown maybe we could have done some analysis on that. There is also a weeks on list which is important for books that have been on more than once.
```{r}
book1_df <- as.data.frame(book1) %>%
  select(
    title = title,
    author = author,
    publisher = publisher,
    rank = rank,
    weeks_on_list = weeks_on_list
)
head(book1_df)
```
Now that we see the formatting of the lists, we can try to extract each of these into the layer above. We can use the unnest function for this which will convert a list in a column into their own column values for each list item. I will also select columns.
```{r}
books_unnested <- books_df %>%
  unnest(books) %>%
    select(
      category = list_name,
      title = title,
      author = author,
      publisher = publisher,
      updated = updated,
      rank = rank,
      weeks_on_list = weeks_on_list
    )
head(books_unnested)
```
Now we have our books in a format that is easier to work with. Something to note is that some categories are updated MONTHLY rather than WEEKLY. Lets take a look at those. \
We can see that the data is formatted such that there is 
```{r}
head(books_unnested %>% 
       filter(
        updated == "MONTHLY"
        )
     )
```
We also are missing the date. The date was in the results that are a layer above this one. I will create a function to get the best sellers of a specific date. \
While making the function to accept the date I notice that the results also have the next date stored in them. Using this we can call the function with the next date repeatedly. \
I decide to make a nested function to perform this. The outer function will contain the date range we want to search through. It will manage the sequential calling of the helper function.
The helper will be responsible for adding data frames to a list, returning the updated list and the previous publication date.


```{r}

get_books_df<- function(date, weeks){
  # Here I define the helper function
  #Takes date in YYYY-MM-DD format(string), returns the given list with the dataframe added, as well as the previous publication date in a list
  get_books_date <- function(date, books_list){ 
    url <- "https://api.nytimes.com/svc/books/v3/lists/overview.json"
    response <- GET(
      url,
      query = list( # We can put parameters here, api-key is needed and we can specify a date.
        'published_date' = date,
        'api-key' = api_key
        )
    )
    data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
    json_results <- fromJSON(data, flatten = TRUE) # 
    books_results <- json_results$results
    # We need the save the published_date from this layer, the api describes this as "When the list was in print."
    publish_date <- books_results$published_date
    # I will also save next and previous publish dates
    next_publish_date <- books_results$next_published_date #UNUSED
    previous_publish_date <- books_results$previous_published_date
    # Now we can create the dataframe
    books_df <- json_results$results$lists
    books_unnested <- books_df %>%
    unnest(books) %>%
      select(
        category = list_name,
        title = title,
        author = author,
        publisher = publisher,
        updated = updated,
        rank = rank,
        weeks_on_list = weeks_on_list
      ) 
    books_unnested$publish_date = publish_date
    books_list[[length(books_list) + 1]] <- books_unnested
    return(list(books_list, previous_publish_date))
  }
  # I need the outer function to create the list that we are adding to.
  books_list <- list()
  # It will create the book list with calls to get_books_date, starting at the date given, for weeks as the number of times we go back.
  date_index = date
  while(weeks > 0){
    books_date_results <- get_books_date(date_index, books_list)
    weeks <- weeks - 1 #Decrease the weeks counter
    books_list <- books_date_results[[1]] # Update book List 
    date_index <- books_date_results[[2]] # Update the index to the previous date
    Sys.sleep(10) # Sleep to prevent rate limit
  }
  #Now we should have a books_list that contains all our dataframes.
  # Return these rows after being combined
  return(bind_rows(books_list))
}  
  
```
I will use this to get the past 6 weeks from today's date. I added a sleep to prevent the rate limit to the function but it will seemingly randomly break if too many weeks are done at once.
```{r}
books_df <- get_books_df("2026-03-29", 6)
```
I believe that having 6 weeks of data will provide practically the same steps in analysis so I will stick with 6 weeks for now.

```{r}
head(books_df)
```
Now that we have a dataframe in a tidy format we can start to perform some analysis.\
My initial question was which authors and publishers are published the most.\
We can group by author and publisher (separately and sum to get the count) to see who was published the most
```{r}
author_count <- books_df %>%
  group_by(author) %>%
  summarize(
    times_published = n()
  ) %>%
  arrange(desc(times_published))
head(author_count)
```
I will make a table to show this
```{r}
author_count %>%
  gt() %>%
      cols_label(
        'author' = "Author",
        'times_published' = "Times published"
      ) %>%
      tab_header(title = md("Count of publications in NYTimes Best Sellers (Past 6 weeks)")) %>%
      tab_options(container.height = 500, container.overflow.y = TRUE) # Set container height and container overflow to add scrolling for large tables.
```
I will make another table for the publishers
```{r}
publisher_count <- books_df %>%
  group_by(publisher) %>%
  summarize(
    times_published = n()
  ) %>%
  arrange(desc(times_published))
#Display in table
publisher_count %>%
  gt() %>%
      cols_label(
        'publisher' = "Publisher",
        'times_published' = "Times published"
      ) %>%
      tab_header(title = md("Count of publications in NYTimes Best Sellers (Past 6 weeks)")) %>%
      tab_options(container.height = 500, container.overflow.y = TRUE) # Set container height and container overflow to add scrolling for large tables.

```


