---
title: "607 Assignment 8 Dylan Gold"
format: pdf
editor: visual
---

## Approach

In this assignment we have to make use of the NYtimes api to answer a question through queries made through the api. The hardest part of this initally was coming up with a research question. I started off by testing out the api after creating an account. I tried using the article search api and found that using the filter did not really help that much for some of my inital examples. I tried restaurants and many of the articles seeming had nothing to do with restaurants.

While doing some inital setting up of the api key in R I decide to use the usethis library to adjust the environment.\
In the console I type the following to save a api key to my environment under "NYT_API_KEY". R may need to be restarted after.\

```         

library(usethis)
edit_r_environ()
```

```{r , message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
```

Now we have our key

```{r}
api_key <- Sys.getenv("NYT_API_KEY")
```

After some testing I found that the nytimes api works with this query list. Other ways of adding it as a header did not seem to work for me but here we get an example of the current hardcover-fiction books

```{r}
url <- "https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json"
response <- GET(
  url,
  query = list('api-key' = api_key) #here we list query parameters,
)
status_code(response)
```

Looking at the data in our environment from this we can see there is a lot of extra information. I will figure out how to extract the data I need at a later point because I am still not certain of my question to ask.\

```{r}
books <- content(response)
head(books$results)
```

I will probably look at the books to answer a question about comparing different types of books or I will use the article search api to look at how many articles are made each year with a certain filter and see how things have increased. For example given some of the recent news about insider trading maybe I will use insider trading as a filter for each year within the past 15 years and see how the occurence of these articles have either decreased or increased over time. I may pick something else as the search filter.\
At this point I at least have the api set up minimally and I will work on parsing it and deciding on my specific question later.\

## Codebase

First I will decide the question I want to answer. Given I have looked at several data sets relating to energy consumption, carbon emissions and other global warming adjacent topics this semester, I will try to answer: How has coverage of climate change changed over the past 10 years in the New York Times.\

I will set the api key up again like I did in my approach

```{r , message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
api_key <- Sys.getenv("NYT_API_KEY")

```

I will start off by testing the article search. It is different than the book request in that we do not need to put the query parameters in the url.\
I believe this is because this has no required parameters. We should be able to add these in the query of the GET function.\
I will test with the query as climate change. I will look in the year 2025

```{r}
url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json"
response <- GET(
  url,
  query = list(
    'begin_date' = 20250101, # Start date 2025 01 01
    'end_date' = 20251231, # End date 2025 12 31
    'q' = "Climate change", #Query parameter
    'api-key' = api_key
    )
)
status_code(response)
```

I was not sure how to extract the data initially. Using ?GET to see the responder object and content function was alright but I was not sure of the type or encoding that I should be using. I used a LLM to help me create the dataframe because of this.  I used chatgpt to help generate the following code.

```{r}
data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
json_data <- fromJSON(data, flatten = TRUE) # 
df <- json_data$response$docs
```

Now that we have a data frame we can try to create a tidy data frame from this that contains the information we need. We are interested in seeing how the coverage of articles with climate change as a topic have increased or decreased on the NY times. We only really need the name and date the articles were published.\
I will start off with selecting columns I want. We can see the columns with names()

```{r}
names(df)
class(df)
```

I will select the headline.main, publication date, url.

```{r}
articles <- df %>% select(
  headline.main, 
  pub_date,
  web_url
)
head(articles)
```

```{r}
nrow(articles)
```

I had tried to adjust the initial request to a larger date range and realize that I am only getting 10 responses. This is because the api will give 10 articles at a time separated by pages. We can rerun the request and creation of the data frame with a different page to see how the articles are different to confirm this. I also put in 0 as the page to see if it matched the previous dataframe created and it does.

```{r}
url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json"
response <- GET(
  url,
  query = list(
    'begin_date' = 20250101, # Start date 2025 01 01
    'end_date' = 20251231, # End date 2025 12 31
    'q' = "Climate change", #Query parameter
    'page' = 100,
    'api-key' = api_key
    )
)
data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
json_data <- fromJSON(data, flatten = TRUE) # 
df <- json_data$response$docs
articles <- df %>% select(
  headline.main, 
  pub_date,
  web_url
)

articles
```

After testing 100 I realize that with our current search we are hitting articles that may not have to do with climate change. I need a more concrete way of searching for climate change articles. I asked an LLM to show me how to filter it by topic as climate change. I had tried to do it the way the llm suggested but at this point it was not giving me working responses.\
I tried to add climate change as a filter aswell with global warming but I was still getting too many results.

```{r}
url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json"
response <- GET(
  url,
  query = list(
    'begin_date' = 20250101, # Start date 2025 01 01
    'end_date' = 20251231, # End date 2025 12 31
    'q' = "Climate change", #Query parameter
    'fq' = 'climate change, global warming', # Added for stricter search
    'page' = 100,
    'api-key' = api_key
    )
)
data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
json_data <- fromJSON(data, flatten = TRUE) # 
df <- json_data$response$docs

articles <- df %>% select(
  headline.main, 
  pub_date,
  web_url
)
articles
```

I decide to change my approach. Rather than make the searches more specific I would attempt to collect more data. rather than by year I will try by month. After trying 100 pages in a 2 month frame I get null values. Not all the values are null, but it is good that the pages at the end are, otherwise we could not collect all the results. I also notice that I get better results if I just filter for climate change as well as the query.

```{r}
response <- GET(
  url,
  query = list(
    'begin_date' = 20250101, # Start date 2025 01 01
    'end_date' = 20250331, # End date 2025 12 31
    'q' = "Climate change",
    'fq' = 'climate change', # Added for stricter search
    'page' = 0,
    'api-key' = api_key
    )
)
data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
json_data <- fromJSON(data, flatten = TRUE) # 
df <- json_data$response$docs

articles <- df %>% select(
  headline.main, 
  pub_date,
  web_url,
  keywords # Added to see if we can use this to help determine which articles are climate change related.
)
articles
```

While this may be a solution for the problem of not having enough room for the articles, I also see how we are getting some articles that do not have to do with climate change.\
I notice that there is a keywords column in the response I will look at this a little to see if climate change is a keyword that is used.

```{r}
articles$keywords
```

This shows us each article has a data frame that shows different types of tags. I now see that Global Warming is a popular keyword. Before I was looking for climate change. I will adjust my search to have climate change as the query and subject = "Global Warming" as a filter.

```{r}
response <- GET(
  url,
  query = list(
    'begin_date' = 20250101, # Start date 2025 01 01
    'end_date' = 20250331, # End date 2025 12 31
    'q' = "Climate change",
    'fq' = 'subject="Global Warming"', # Added for stricter search
    'page' = 7,
    'api-key' = api_key
    )
)
data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
json_data <- fromJSON(data, flatten = TRUE) # 
df <- json_data$response$docs

articles <- df %>% select(
  headline.main, 
  pub_date,
  web_url,
  keywords # Added to see if we can use this to help determine which articles are climate change related.
)
articles
```

I will now try to obtain the past 10 years of articles. I will create a function that will perform the query and pass it different ranges for each month with a sequence.

```{r}
get_response <- function(start, end) {
  # List of dataframes for each page to fill, initially empty
  article_list <- list()
  # Boolean to keep track if we have more articles
  more_articles <- TRUE 
  # Page number to index
  page_number <- 0
  #I will use a while loop, this will let us get the pages with material but stop once we get the pages
  while(more_articles){
    response <- GET(
      url,
      query = list(
        'begin_date' = start, # Start date 2025 01 01
        'end_date' = end, # End date 2025 12 31
        'q' = "Climate change",
        'fq' = 'subject="Global Warming"', # Added for stricter search
        'page' = page_number,
        'api-key' = api_key
        )
    )
    
    data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
    json_data <- fromJSON(data, flatten = TRUE) # 
    df <- json_data$response$docs
    
    #Here check if the dataframe is null. If the page is empty it will be null. and we stop searching
    if (!is.null(df)){
    article <- df %>% select( #Single article
      headline.main, 
      pub_date,
      web_url
    )
    #update the list
    article_list[[length(article_list) + 1]] <- article
    #Increment page_number
    page_number <- page_number + 1
    }
    else{ # Set more_articles to false if we have an empty dataframe
      more_articles <- FALSE
    }
  }
  
  return(bind_rows(article_list)) # Returns the list combined into a single dataframe
}
```

I will test the function. It should get all pages for a range given.

```{r}
article_pages <- get_response(20250101, 20250331)
```

```{r}
article_pages2 <- get_response(20200101, 20200331)
```

We can see if we got more than the initial 10 pages and it seems we did.

```{r}
head(article_pages)
nrow(article_pages)
```

At this point I see that I am stuck at 50 observations in my dataframes. This means I am only getting 5 pages. In the first example I should be getting more than 5 pages. It is likely because I am being rate limited somehow. I decide I need to take a different approach. I decide to leave what I wrote so far in to show my process. It has also given me a better understanding of the data.\
\## Codebase Attempt #2 There is another api url from the new york times that gives us all the articles written in a certain month and year. I will use this along with the keyword to see the trend of published articles with the keyword subject "Global Warming".\
I will create a function that gets all these articles for a certain month an year

```{r}

url <- "https://api.nytimes.com/svc/archive/v1/2025/1.json "
response <- GET(
  url,
  query = list('api-key' = api_key) 
)
status_code(response)
data <- content(response, as = "text", encoding = "UTF-8") # Get data from request object
json_data <- fromJSON(data, flatten = TRUE) # 
df <- json_data$response$docs

articles <- df %>% select(
  headline.main, 
  pub_date,
  web_url
)
articles
```
